---
title: "DWVAssignment5.qmd"
format: html
---

## 2. Re-code 5-category ideology

```{r}
library(tidyverse)
library(haven)
ces <- read_dta("~/DataWrangling/Week5/CES 2006-2024 Cumulative/cumulative_2006-2024.dta")
ces2024 <- ces |> filter(year == 2024)
print_labels(ces2024$ideo5)

ces2024 <- ces2024 |>
  mutate(ideo5_recode = case_when(
    ideo5 %in% 1:2 ~ -1,
    ideo5 %in% 3 ~ 0,
    ideo5 %in% 4:5 ~ 1,
    ideo5 == 6    ~ NA
  ))
```

## 3. Re-code presidential vote

```{r}
print_labels(ces2024$voted_pres_24)
ces2024 <- ces2024 |>
  mutate(voted_pres_24_recode = case_when(
    voted_pres_24 %in% 3:8 ~ NA,
    voted_pres_24 == 1 ~ 1, 
     voted_pres_24 == 2 ~ 2
  ))
```

## 4. Table

```{r}
ces2024 <- ces2024 |>
mutate(ideo5_recode2 = case_when(
  ideo5 %in% 1:2 ~ 1,
  ideo5 == 3 ~ NA,
  ideo5 %in% 4:5 ~ 2,
  ideo5 == 6 ~ NA))
    
VoteShare = table(Vote = ces2024$voted_pres_24_recode, Ideology = ces2024$ideo5_recode2)
VoteShare
VoteShare1 = prop.table(VoteShare, 2)
VoteShare1


```

2.7% of liberal voters voted for Donald Trump, while 7.2% of conservative voters voted for Kamala Harris.

## 5. Rent or own

```{r}
print_labels(ces2024$ownhome)

ces2024 <- ces2024 |>
  mutate(ownhome_recode = case_when(
    ownhome == 1 ~ 1, 
    ownhome == 2 ~ 2,
    ownhome == 3 ~ NA
  ))

```

## 6. Rent or own table

```{r}
print_labels(ces2024$ownhome)

ces2024 <- ces2024 |>
  mutate(ownhome_recode = case_when(
    ownhome == 1 ~ 1, 
    ownhome == 2 ~ 2,
    ownhome == 3 ~ NA
  ))

OwnHomeVote = table(Vote = ces2024$voted_pres_24_recode, HomeOwnership = ces2024$ownhome_recode)
OwnHomeVote
OwnHomeVote1 = prop.table(OwnHomeVote, 2)
OwnHomeVote1


```

48.4% of homeowners who voted for one of the two major candidates voted for Trump, while 51.6% of homeowners voted for Harris. 64.8% of voters who rent voted for Harris, while 35.2% voted for Trump.

## 7. Education Recode 
```{r}
print_labels(ces2024$educ)
ces2024 <- ces2024 |>
  mutate(educ_recode = case_when(
    educ %in% 1:2 ~ 1, 
    educ %in% 3:4 ~ 2,
    educ == 5 ~ 3,
    educ == 6 ~ 4
  ))

```
## 8. Education and voting table 
```{r}
EducationVote = table(Vote = ces2024$voted_pres_24_recode, EducationLevel = ces2024$educ_recode)
EducationVote
EducationVote1 = prop.table(EducationVote, 2)
EducationVote1

```
The data demonstrate that as education level rises, the proportion of individuals who voted for Harris increases, and the proportion who voted for Trump decreases. Notably, those with a post-graduate education are significantly more likely to have voted for Harris than any other group. 
